### 2020-05-24 intro.tex
updated outline

We are not, of course, the first to integrate direct manipulation user interfaces
into general-purpose programming systems.
Prior work on projectional editing
and active code completion, summarized in Sec. \ref{sec:related-work}\todo{sections}{}, 
has also considered the problem of entering expressions 
of certain types, like \li{Color}, 
using specialized GUIs integrated into an editor. 
The prior work most closely related to this paper is the {Graphite} system for Java \cite{Graphite}.
Graphite allows library providers to associate GUIs, called \emph{palettes}, with types (using Java's class annotations). 
Wherever an expression of a type equipped with a palette is needed, 
i.e. wherever there is a \emph{hole} of that type in the program, 
the programming environment offers the programmer the option, via the code completion menu, 
to activate the palette. 
Once the interaction is finished, it is the palette's job to generate an 
appropriate Java expression to fill the hole. 
Figure~\ref{fig:color}(a)\todo{use subfigure?}{}, adapted from this prior work, demonstrates a simple color palette invoked using Graphite.
When the user presses the \li{Enter} key, the Java code \li{new Color(0, 0, 128)}\todo{correct code}{} is inserted at the cursor (not shown) and the palette disappears.

@@

  \item \textbf{Parameterization}: It is possible to define parameterized families of livelits. 
  For example, \li{\$slider} in Fig.~\ref{fig:color}(b) is parameterized by the lower and upper bound. 
  Parameters behave like splices, differing in that livelit applications can be abbreviated, and abbreviations
  support partial application. For example, it is possible to define a livelit named \li{\$percentage} 
  that operates as a \li{\$slider} from \li{0} to \li{100} by scoped abbreviation:
  \begin{lstlisting}[numbers=none]
  let livelit $percentage = $slider(0, 100) in ...
  \end{lstlisting}
  Parameterization also underlies the hygiene mechanism, as we will discuss in Sec. 4\todo{sec}{}.

@@

The remainder of this paper is organized as follows. We begin in Sec.~2 with two case studies 
involving several non-trivial livelits, focusing on the livelit client's perspective. 
We then consider the livelit provider's perspective in Sec.~3.
In Sec.~4, we provide a formal account of the mechanism as a typed lambda calculus. 
This calculus formally specifies the essential aspects of livelit definition, representation, expansion, 
and run-time closure collection independent of the particularities of our implementation, 
and establishes essential type safety and hygiene properties. 
In Sec.~5, we describe some extensions of the core mechanism\todo{what are the extensions?}{}.
In Sec.~6, we outline our implementation of livelits within Hazel.
In Sec.~7, we situate livelits relative to other related work. 
Finally, we conclude in Sec.~8 with a discussion of some present limitations and future work.\todo{section refs}{}

>>>>

We are not, of course, the first to integrate direct manipulation user interfaces
into general-purpose programming systems.
Prior work on projectional editing
and active code completion, summarized in Sec. \ref{sec:related-work}, 
has also considered the problem of entering expressions 
of certain types, like \li{Color}, 
using specialized GUIs integrated into an editor. 
The prior work most closely related to this paper is the {Graphite} system for Java \cite{Graphite}.
Graphite allows library providers to associate GUIs, called \emph{palettes}, with types (using Java's class annotations). 
Wherever an expression of a type equipped with a palette is needed, 
i.e. wherever there is a \emph{hole} of that type in the program, 
the programming environment offers the programmer the option, via the code completion menu, 
to activate the palette. 
Once the interaction is finished, it is the palette's job to generate an 
appropriate Java expression to fill the hole. 
Figure~\ref{fig:color}(a)\todo{use subfigure?}{}, adapted from this prior work, demonstrates a simple color palette invoked using Graphite.
When the user presses the \li{Enter} key, the Java code \li{new Color(0, 0, 128)}\todo{correct code}{} is inserted at the cursor (not shown) and the palette disappears.

@@

  \item \textbf{Parameterization}: It is possible to define parameterized families of livelits. 
  For example, \li{\$slider} in Fig.~\ref{fig:color}(b) is parameterized by the lower and upper bound. 
  Parameters behave like splices, differing in that livelit applications can be abbreviated, and abbreviations
  support partial application. For example, it is possible to define a livelit named \li{\$percentage} 
  that operates as a \li{\$slider} from \li{0} to \li{100} by scoped abbreviation:
  \begin{lstlisting}[numbers=none]
  let livelit $percentage = $slider(0, 100) in ...
  \end{lstlisting}
  Parameterization also underlies the hygiene mechanism, as we will discuss in Sec.~\ref{sec:livelit-definitions}.

@@

The remainder of this paper is organized as follows. We begin in Sec.~\ref{sec:case-studies} with two case studies 
involving several non-trivial livelits, focusing on the livelit client's perspective. 
In Sec.~\ref{sec:livelit-definitions}, we consider the livelit provider's perspective by detailing the livelit
definition mechanism by example. 
In Sec.~\ref{sec:livelit-calculus}, we provide a formal account of the livelit mechanism as a typed lambda calculus. 
This calculus formally specifies the essential aspects of livelit definition, representation, expansion, 
and run-time closure collection independent of the particularities of GUI framework and other implementation details. 
This calculus and the key metatheoretic properties has been mechanized in Agda.
In Sec.~\ref{sec:extensions}, we describe some extensions of the core calculus\todo{what are the extensions?}{}.
In Sec.~\ref{sec:implementation}, we provide a more detailed account of our implementation of livelits within Hazel, 
and discuss several implementation details that would need to be considered by other implementations.  
In Sec.~\ref{sec:related-work}, we compare livelits to related work. 
Finally, we conclude in Sec.~\ref{sec:discussion} with a discussion of some present limitations and future work.
